---
title: "model-object"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{model-object}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(hyperion)
```
```{r}
model_dir <- system.file("extdata/models/onecmt", package = "hyperion")
lookup <- system.file("lookup.toml", package = "hyperion")
```

# Model Object

create the model object by giving the path to the model file, and optionally the lookup toml path.

```{r}
run003 <- Model(file.path(model_dir, "run003.mod"), lookup)
```
You still get the erros with incorrect paths

```{r, error=TRUE}
print(
  try(
    run003_error_dir <- Model(file.path(model_dir, "run003")),
  )
)
```	

```{r, error=TRUE}
print(
	try(
		run003_error_ouput <- Model(file.path(model_dir, "run003/run003.mod")),
	)
)
```
the `hyperion_nonmem_model` object is the `@model` property

```{r}
run003@model
```

# all get_* functions are just `getter` properties

`get_model_summary(mod)` is the `@summary` property: 
```{r}
run003@summary
```

`get_model_parameter_info(mod)` is the `@info` property:
```{r}
run003@info
```
`audit_parameter_info(info)` is the `@info_audit` property:
```{r}
run003@info_audit
```
`@info` is cached such that updates persist
```{r}
run003@info@sigma$`SIGMA(2,2)`@parameterization <- "AddErr"
run003@info_audit
```
`get_model_lineage` is the `@lineage` property:
```{r}
run003@lineage
```
you can get `@ancestors` and `@descendants`
```{r}
run003@ancestors
```
```{r}
run003@descendants
```
You can get `@run_info`

```{r}
run003@run_info
```
`@gradients`, `@eps_shrinkage`, `@eta_shrinkage`
```{r}
run003@gradients
```

```{r}
run003@eta_shrinkage
```

```{r}
run003@eps_shrinkage
```

# Parameter Table creation

`@parameters` can be used for parameter tables
```{r}
spec <- TableSpec(
  display_transforms = list(omega = c("cv")),
  sections = section_rules(
    kind == "THETA" ~ "Structural model parameters",
    kind == "OMEGA" & diagonal ~ "Interindividual variance parameters",
    kind == "OMEGA" & !diagonal ~ "Interindividual covariance parameters",
    kind == "SIGMA" ~ "Residual error",
    TRUE ~ "Other"
  ),
  name_source = "display",
  title = paste(run003@model_name, "Parameters")
)

run003@parameters |>
  apply_table_spec(spec, run003@info) |>
  add_summary_info(run003@summary) |>
  make_parameter_table()
```
# Comparison table

```{r}
comp_spec <- spec
comp_spec@title <- "Model Comparison"
comp_spec@drop_columns <- c("rse", "shrinkage", "variability")

run002 <- Model(file.path(model_dir, "run002.mod"), lookup)

run002@parameters |>
  apply_table_spec(comp_spec, run002@info) |>
  add_summary_info(run002@summary) |>
	compare_with(
		run003@parameters |>
			apply_table_spec(comp_spec, run003@info) |>
			add_summary_info(run003@summary),
		labels = c(run002@model_name, run003@model_name)
	) |>
	add_model_lineage(run002@lineage) |>
  make_comparison_table()
```

# Summary Table

```{r}
sum_spec <- SummarySpec()

run003@lineage |>
	apply_summary_spec(sum_spec) |>
	make_summary_table()
```

