---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hyperion

<!-- badges: start -->
[![R-CMD-check](https://github.com/A2-ai/hyperion/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/A2-ai/hyperion/actions/workflows/R-CMD-check.yaml)
[![Pharos Dependency](https://github.com/A2-ai/hyperion/actions/workflows/pharos-dependency-check.yaml/badge.svg)](https://github.com/A2-ai/hyperion/actions/workflows/pharos-dependency-check.yaml)
<!-- badges: end -->

Hyperion is a companion R packge to the cli tool [pharos](https://github.com/a2-ai/pharos) for managing and running pharamceutical models directly from R. NONMEM is the first supported modelling software.

## Installation

You can install the development version of hyperion from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("A2-ai/hyperion")
```
## Getting started

To initialize hyperion/pharos use the `hyperion::init()` cfunction to create a `pharos.toml` configuration file 

```{r example}
library(hyperion)
test_data_dir <- system.file("extdata", "test_data", package = "hyperion")

if (!file.exists("pharos.toml")) {
  hyperion::init(".")
}
```
The `pharos.toml` file contains several configuration options for NONMEM and pharos. You can see more detailed explanations [from pharos](https://github.com/A2-ai/pharos?tab=readme-ov-file#nonmem)

## Checking a model

You can check a model for correct compilation before submitting to catch any data path issues, or syntax errors within the control stream with:
```{r check}
check_model(file.path(test_data_dir, "models", "onecmt", "run002a.mod")) |>
  cat()
```

```{r}
check_model(file.path(test_data_dir, "models", "onecmt", "run004.mod")) |>
  cat()
```


## Viewing a model object

Hyperion can read .mod files to give an overview of the mod file with:
```{r model, results='asis'}
read_model(file.path(test_data_dir, "models", "onecmt", "run002.mod"))
```

## Running a model

There is no current support from hyperion to run a model, but SLURM job submission will be coming soon.

## Model summary

After running a model you can view run details and final estimates with:
```{r summary, results='asis'}
get_model_summary(file.path(test_data_dir, "models", "onecmt", "run002"))
```

### Parameter Table
```{r parameter-table, results='asis'}
spec <- TableSpec(
  display_transforms = list(omega = c("cv")),
  sections = section_rules(
    kind == "THETA" ~ "Structural model parameters",
    kind == "OMEGA" & diagonal ~ "Interindividual variance parameters",
    kind == "OMEGA" & !diagonal ~ "Interindividual covariance parameters",
    kind == "SIGMA" ~ "Residual variance",
    TRUE ~ "Other"
  )
)

mod_sum <- get_model_summary(file.path(test_data_dir, "models", "onecmt", "run003"))
info <- get_model_parameter_info(file.path(test_data_dir, "models", "onecmt", "run003"))

get_parameters(file.path(test_data_dir, "models", "onecmt", "run003")) |>
  apply_table_spec(spec, info) |>
  add_summary_info(mod_sum) |>
  make_parameter_table() |>
  gt::as_raw_html()
```

## Copying a model

You can copy a model to a new control stream and alter the initial estimates of the new model. This will create a new mod file and a `*_metadata.json` file that contains the description and which model it is based on. 

```{r copy}
copy_model(
  from = file.path(test_data_dir, "models", "onecmt", "run002.mod"),
  to = file.path(test_data_dir, "models", "onecmt", "run002a.mod"),
  update = "all", # sets initial estimates of `to` with final estimates of `from`
  jitter = 0.1, # jitters run002a initial estimates by 10%
  description = "Some description about what makes run002a different",
  overwrite = TRUE,
  seed = 804831
)
```
## Model lineage

If you use hyperion to copy models you can extract the model lineage with
```{r lineage, results='asis'}
get_model_lineage(file.path(test_data_dir, "models", "onecmt"))
```
