---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hyperion

<!-- badges: start -->
[![R-CMD-check](https://github.com/A2-ai/hyperion/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/A2-ai/hyperion/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

Hyperion is a companion R packge to the cli tool [pharos](https://github.com/a2-ai/pharos) for managing and running pharamceutical models directly from R. NONMEM is the first supported modelling software.

## Installation

You can install the development version of hyperion from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("A2-ai/hyperion")
```
## Getting started

To initialize hyperion/pharos use the `hyperion::init()` cfunction to create a `pharos.toml` configuration file 

```{r example}
library(hyperion)

if (!file.exists("pharos.toml")) {
  hyperion::init(".")
}
```
The `pharos.toml` file contains several configuration options for NONMEM and pharos. You can see more detailed explanations [from pharos](https://github.com/A2-ai/pharos?tab=readme-ov-file#nonmem)

## Checking a model

You can check a model for correct compilation before submitting to catch any data path issues, or syntax errors within the control stream with:
```{r check}
hyperion::check_model("vignettes/test_data/models/onecmt/run001.mod")
```
## Viewing a model object

Hyperion can read .mod files to give an overview of the mod file with:
```{r model}
read_model("vignettes/test_data/models/onecmt/run002.mod")
```

## Running a model

There is no current support from hyperion to run a model, but SLURM job submission will be coming soon.

## Model summary

After running a model you can view run details and final estimates with:
```{r summary}
get_model_summary("vignettes/test_data/models/onecmt/run002")
```

## Copying a model

You can copy a model to a new control stream and alter the initial estimates of the new model. This will create a new mod file and a `*_metadata.json` file that contains the description and which model it is based on. 

```{r copy}
copy_model(
  from = "vignettes/test_data/models/onecmt/run002.mod", 
  to = "vignettes/test_data/models/onecmt/run002a.mod",
  update = "all", #sets initial estimates of `to` with final estimates of `from` 
  jitter = 0.1, #jitters run002a initial estimates by 10%
  description = "Some description about what makes run002a different",
  overwrite = TRUE,
  seed = 804
)
```
## Model lineage

If you use hyperion to copy models you can extract the model lineage with
```{r lineage}
get_model_lineage("vignettes/test_data/models/onecmt")
```
